---
title: "Lending_club_sample"
author: "Snehal"
date: "12/3/2018"
output: html_document
---

```{r load, echo=FALSE, warning=FALSE, message=FALSE}
library(choroplethr)
library(choroplethrMaps)
library(DescTools)
library(dplyr)
library(ggplot2)
library(readr)

options(scipen=999)
loanbook <- read_csv("loan.csv")

# load the state names
data(state.regions)
# merge the loan book with the state names
loanbook <- merge(loanbook, state.regions, by.x = "addr_state", by.y = "abb")

```

## Data overview

After loading the entire data set I'd like to check its structure and summarise it. 

```{r, echo = TRUE}
# print dimentions
dim(loanbook)

# print column names
colnames(loanbook)
```

I'm not printing these results for conciseness:

```{r, eval=FALSE}
str(loanbook)
summary(loanbook)
```

## Missing variables

I've also noticed that several key fields described in the data dictionary were missing from the loan book:

```{r, echo=TRUE}
library(readxl)

dataDictionary <- read_excel("LCDataDictionary.xlsx")

# fields available in the data dictionary
dd_names <- as.character(na.omit(dataDictionary$LoanStatNew))

# fields available in the loan book
loanbook_names <- names(loanbook)

# show the fields described in data dictionary but not in the loan book
setdiff(dd_names, loanbook_names)
```

## Loan amounts

I focused on several variables and I plotted them using DescTools package.
Here I created the density plot, box plot, and empirical distribution function plot.
```{r, echo=TRUE}
Desc(loanbook$int_rate, main = "interest rate distribution", plotit = TRUE)
```

Here's how the loan book was growing:
```{r, echo=TRUE}
loanbook$issue_d <- as.Date(gsub("^", "01-", loanbook$issue_d), format="%d-%b-%Y")

amnt_df <- loanbook %>% 
  select(issue_d, loan_amnt) %>% 
  group_by(issue_d) %>% 
  summarise(Amount = sum(loan_amnt))

ts_amnt <- ggplot(amnt_df, 
                  aes(x = issue_d, y = Amount))
ts_amnt + geom_line() + xlab("Date issued")
```

## Loan statuses

The loanbook consists of loans in various statuses so I started with exploring their frequency.
```{r, echo=TRUE}
Desc(loanbook$loan_status, plotit = T)
```

Then I checked the distribution of loan amounts by status.
```{r, echo=TRUE}
box_status <- ggplot(loanbook, aes(loan_status, loan_amnt))
box_status + geom_boxplot(aes(fill = loan_status)) +
  theme(axis.text.x = element_blank()) +
  labs(list(
    title = "Loan amount by status",
    x = "Status",
    y = "Amount"))  
```

Here's how the value of loans of different grades was changing over time
```{r, echo=TRUE}
amnt_df_grade <- loanbook %>% 
  select(issue_d, loan_amnt, grade) %>% 
  group_by(issue_d, grade) %>% 
  summarise(Amount = sum(loan_amnt))

ts_amnt_grade <- ggplot(amnt_df_grade, 
                  aes(x = issue_d, y = Amount))
ts_amnt_grade + geom_area(aes(fill=grade)) + xlab("Date issued")
```

## Maps

In order to create maps of loans I initially joined the loan book 
with the names of states and then I created a data frame required by choroplethr.

### Loans by value
```{r, echo=TRUE}
state_wise_loan_amnt <-
loanbook %>% group_by(region) %>%
  summarise(value = sum(loan_amnt, na.rm=TRUE))

state_choropleth(state_by_value, title = "Loan amount by State")
```

### Loans by volume
```{r, echo=TRUE}
state_wise_no_ofloan <-
loanbook %>% group_by(region) %>%
  summarise(value = n())

state_choropleth(state_by_volume, title = "No of loans by State")
```

## Loan reasons

### What's the reason for taking a loan with LendingClub?

```{r, echo=TRUE}
Desc(loanbook$purpose, main = "Loan purposes", plotit = TRUE)
Desc(loanbook$title, main = "Loan titles", plotit = TRUE)
```

## Word cloud

Word cloud gives a good overview of the loan titles (gives by borrowers). 
This information should explain what kind of loans are being funded by LendingClub.

In order to create a word cloud, first I loaded the necessary libraries, 
then I preprocessed the 'title' column by removing punctuation and transforming it to lower case.
This analysis was run on the first 10000 rows to speed up the process.

```{r, echo=TRUE, warning=FALSE, message=FALSE}
library(tm)
library(RColorBrewer)
library(wordcloud)
library(corpus)
v<-data.frame(head(loanbook[,23], n=10000))
doc_id <- rownames(v)
v <- cbind(id=doc_id, v)
names(v) <- c("doc_id", "text")


loan_descriptions.corpus <- Corpus(DataframeSource(v))
loan_descriptions.corpus <- tm_map(loan_descriptions.corpus, removePunctuation)
loan_descriptions.corpus <- tm_map(loan_descriptions.corpus, content_transformer(tolower))

wordcloud(loan_descriptions.corpus,
          max.words = 100,
          random.order=FALSE, 
          rot.per=0.30, 
          use.r.layout=FALSE, 
          colors=brewer.pal(8, "Paired"))
```

## Loan grades

Here is the overview of the occurrence of loans of different grades:
```{r, echo=TRUE}
Desc(loanbook$grade, main = "Loan grades", plotit = TRUE)
```

The last step (so far) was checking whether the interest rates are dependent on the loan grade.
```{r, echo=TRUE}
Desc(int_rate ~ grade, loanbook, digits = 1, main = "Interest rate by grade", plotit = TRUE)
```

Unsurprisingly, the higher the grade (more risky loan), the higher the interest rates.

## Default prediction

There are several statuses which indicate that loans are not performing well. 
I put them into two groups.

```{r, echo=TRUE, warning=FALSE, message=FALSE}
# 'bad' statuses
bad_indicators <- c("Charged Off ",
                    "Default",
                    "Does not meet the credit policy. Status:Charged Off",
                    "In Grace Period", 
                    "Default Receiver", 
                    "Late (16-30 days)",
                    "Late (31-120 days)")

# assign certain statuses to a 'bad' ('0') group
loanbook$is_bad <- ifelse(loanbook$loan_status %in% bad_indicators, 0,
                          ifelse(loanbook$loan_status=="", NA, 1)
                          )
```

Then I wanted to check whether there is a difference between the 'bad' and 'good' loans for numeric variables.

```{r, echo=TRUE, warning=FALSE, message=FALSE}
# figure out which columns are numeric so that we can look at the distribution
numeric_cols <- sapply(loanbook, is.numeric)

# turn the data into long format
library(reshape2)
loanbook.lng <- melt(loanbook[,numeric_cols], id="is_bad")

# plot the distribution for 'bad' and 'good' for each numeric variable
p <- ggplot(aes(x = value, group = is_bad, colour = factor(is_bad)), 
            data = loanbook.lng)

# create the plot to check if there are any good variables that can be used in predictive models
p + geom_density() +
  facet_wrap(~variable, scales="free")
p  +
  facet_wrap(~variable, scales="free")


```

The plots above can be the basis for chosing variables for predictive models (note: scales need fixing).

It seems like Annual Income and Interest Rate are two variables that can be good predictors of how loans will behave.
We can have a closer look at particular cases:

```{r, echo=TRUE, warning=FALSE, message=FALSE}
library(DT)

loanbook %>% 
  filter(is_bad == '0') %>% 
  select(annual_inc, int_rate, loan_status) %>% 
  datatable(., options = list(pageLength = 10))
```
```{r}

loan = loanbook %>%
        select(loan_status , loan_amnt , int_rate , grade , emp_length , home_ownership , 
               annual_inc , term,is_bad)



loan2 = loan %>%
        filter(!is.na(annual_inc) , 
               !(home_ownership %in% c('NONE' , 'ANY')) , 
               emp_length != 'n/a')

## Deriving Feature
loan2 = loan2 %>%
        mutate(loan_outcome = ifelse(loan_status %in% c('Charged Off' , 
                                                        'Default' ,
                                                        "Does not meet the credit policy. Status:Charged Off",
                                                        "In Grace Period", 
                                                        "Default Receiver", 
                                                        "Late (16-30 days)",
                                                        "Late (31-120 days)") , 1,
                                                         ifelse(loan_status == 'Fully Paid' , 0 ,2)
                                                       ))
loan2 = loan2 %>%
        select(-loan_status) %>%
        filter(loan_outcome %in% c(0 , 1))


idx = sample(dim(loan2)[1] , 0.75*dim(loan2)[1] , replace = F)
trainset = loan2[idx , ]
testset = loan2[-idx , ]

train= subset(loan2,loan_outcome==0)
train=train[1:47962,]

train2= subset(loan2,loan_outcome==1)
train=rbind(train,train2)
train=subset(train,select=c(-is_bad))
trainset=subset(trainset,select=c(-is_bad))

testset=subset(testset,select=c(-is_bad))

## Creation of Model


glm.model  = glm(loan_outcome ~ annual_inc+log(annual_inc)+int_rate, trainset, family = binomial(link = 'logit'))

preds = predict(glm.model , testset , type = 'response')
predicted_glm=ifelse(preds>0.65,1,0)
testset$predicted=predicted_glm
testset$predicted=as.factor(testset$predicted)
testset$loan_outcome=as.factor(testset$loan_outcome)

preds=as.factor(preds)
preds=data.frame(preds)
plot(preds)
data_test=data.frame(testset$loan_outcome)
cm <- confusionMatrix(data=testset$predicted, reference=testset$loan_outcome)
cm
summary(predicted_glm)
##Plotting the density plot for GLM
ggplot(data.frame(preds) , aes(preds)) + 
        geom_density(fill = 'lightblue' , alpha = 0.4) +
        labs(x = 'Predicted Probabilities on test set')

library(rpart)
library(rpart.plot)

## Analysis of Grade as a output feature
#loan$lq <- ifelse(loan$grade=='A',"GOOD","BAD")

fit.lm<-rpart(loan_outcome~annual_inc+int_rate,data=train)


rpart.plot(fit.lm)
library(caret)
library(ggplot2)
require(caret)
summary(loan$loan_outcome)
pred_rpart = predict(fit.lm , testset)
predicted_rpart=ifelse(pred_rpart>0.6,1,0)
testset$pred_rpart=predicted_rpart
testset$pred_rpart=as.factor(testset$pred_rpart)
summary(loan$loan_outcome)
summary(pred_rpart)

cm_rpart <- confusionMatrix(data=testset$pred_rpart, reference=testset$loan_outcome)
as.data.frame(table(trainset$loan_outcome))
as.data.frame(table(predicted_rpart))
## SVMMM

library(e1071)
fit.svm<-svm(loan_outcome~.,data=train)
pred_svm1 = predict(fit.svm , testset)
pred_svm=ifelse(pred_svm1>0.8,1,0)
testset$pred_svm=pred_svm
testset$pred_svm=as.factor(testset$pred_svm)
summary(pred_svm1)

as.data.frame(table(pred_svm))

cm_svm <- confusionMatrix(data=testset$pred_svm, reference=testset$loan_outcome)


##Plot the probability density through Decision trees
ggplot(data.frame(pred_rpart) , aes(pred_rpart)) + 
        geom_density(fill = 'lightblue' , alpha = 0.4) +
        labs(x = 'Predicted Probabilities on test set')

library(party)

## Accuracy of different models 
SVM<-round(cm_svm$overall[1],2)
dec_tree<-round(cm_rpart$overall[1],2)
glm<-round(cm$overall[1],2)

q<-data.frame(glm,SVM,dec_tree)

q


```

